### 2019-03-28 report/chapter02.Rnw
Update 28.03

<<echo=FALSE>>=
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'thesis/code')
source(file.path(PATH, 'prepare.R'))

@@

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8.5, fig.height=3,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

@@

\subsection{Structure and Content}
The dataset consists of \Sexpr{length(unique(madata$file.nr))} systematic reviews from the Cochrane Library with \Sexpr{length(unique(madata$study.name))} studies.
Each study provides data of (multiple) comparisons of clinical interventions. 
% It can not be ruled out that some comparisons are retrospective, e.g. from observational studies. 
In Table \ref{barbiturate.row}, two comparisons from a systematic review about effects of barbiturates are shown as they are given in the dataset. As can be seen, the comparison is further specified by the variables in the columns. One row of the dataset is one comparison.

<<echo=FALSE, results = 'asis'>>=
tp1 <- arrange(filter(madata, file.nr == 21) %>% 
          select(Study = study.name, Comparison_type = comparison.name, Outcome = outcome.name, 
                 Events = events1, Total = total1, Events_c = events2, Total_c = total2) %>% 
            slice(c(1,3)), Study)
print(xtable(tp1, label = "barbiturate.row", caption = "Example of two comparisons as given in the dataset. Events denotes the count of events in the treatment group while Events c the count of events in the group compared to. Further descriptive variables have been ommitted"), include.rownames = F, size = "scriptsize")
@

@@

<<echo=FALSE, results='asis'>>=
tp1 <- arrange(filter(madata, file.nr == 21) %>% 
          select(Study = study.name, Comparison = comparison.name,Outcome = outcome.name), Study)
tp1[tp1$Study == "P\303\251rez-B\303\241rcena 2008", 1] = "Perez-Barcena 2008"
print(xtable(tp1, label = "barbiturates", caption = "Barbiturate and head injury review. In the columns, study names, comparison types and outcome measure of the comparisons are given"), include.rownames = F, size = "footnotesize")
@

@@

\vspace{0mm}
Having provided a rough overview over the dataset, now, some more specific information is provided. The dataset consists of \Sexpr{dim(madata)[1]} comparisons and has \Sexpr{dim(madata)[2]} variables that specify the comparisons. The abundance of some missing values in the dataset is given in Table \ref{missing}. For variables as research subject, outcome and subgroup name and event counts there are no missing values. The relative abundance of missing values is very low except for study years.

<<echo=FALSE>>=
missing.mean <- madata %>% filter(!is.na(sd1) & sd1 > 0) %>% filter(!is.na(sd2) & sd2 > 0) %>% 
  summarise(total = length(sd1), na  = sum(is.na(mean1) | is.na(mean1))) %>% select(na)

missing.sd <- madata %>% filter(!is.na(mean1) & !is.na(mean2)) %>% filter(mean1 != 0 & mean2 != 0) %>%
  summarise(total = length(mean1), na.sd  = sum(is.na(sd1) | is.na(sd1)), null.sd = length(sd1[sd1 <= 0 | sd2 <= 0])) %>% 
  mutate(miss = na.sd + null.sd) %>% select(miss)

missing.ssize <- madata %>% summarize(total = length(total1), null.ss = length(total1[total1 <= 0 | total2 <= 0])) %>% 
  select(null.ss)

missing.effect <- sum(is.na(madata$effect))

missing.year <- sum(is.na(madata$study.year)) + length(madata$study.year[madata$study.year == 0])

# missing.outcome <- sum(is.na(madata$effect))

# null.events <- madata %>% filter(is.na(mean1) | mean1 == 0) %>% filter(is.na(mean2) | mean2 == 0) %>% 
#   filter(is.na(sd1) | sd1 <= 0) %>% filter(is.na(sd2) | sd2 <= 0) %>% 
#   summarise(null.ev = length(total1[events1 <= 0 & events2 <= 0]), total = length(events1)) %>% 
#   mutate(f = null.ev/total) %>% select(f)
# 
# null.events.single <- madata %>% filter(is.na(mean1) | mean1 == 0) %>% filter(is.na(mean2) | mean2 == 0) %>% 
#   filter(is.na(sd1) | sd1 <= 0) %>% filter(is.na(sd2) | sd2 <= 0) %>% 
#   summarise(null.ev = length(total1[events1 <= 0 | events2 <= 0]), total = length(events1)) %>% 
#   mutate(f = null.ev/total) %>% select(f)
# 
# missing.events <- madata %>% filter(is.na(mean1) | mean1 == 0) %>% filter(is.na(mean2) | mean2 == 0) %>% 
#   filter(is.na(sd1) | sd1 <= 0) %>% filter(is.na(sd2) | sd2 <= 0) %>% 
#   summarise(na.ev = sum(is.na(events1) | is.na(events2)), total = length(events1)) %>% 
#   mutate(f = na.ev/total) %>% select(f)
@

@@

More properties of the reviews, the studies and the comparisons in the dataset will be provided on the following pages. The publication dates of the studies included in the dataset are shown in Figure \ref{study.years}.

\begin{figure}
<<echo=FALSE>>=
madata %>% filter(study.year < 2019 & study.year > 1920) %>% ggplot(aes(x = study.year)) + geom_histogram(col = "gray15", fill = "dodgerblue") +
  theme_bw() + xlab("Year") + ylab("Study frequency")
@
\caption{Frequencies of study publication years in the dataset. 44655 were excluded due to likely wrong indications}
\label{study.years}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
yodata %>% group_by(outcome.measure) %>% count() %>% arrange(desc(n)) %>% ungroup() %>% filter(row_number() < 9) %>% 
  ggplot(aes(x = outcome.measure, y = n)) +
  geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() + ylab("Comparison frequency") + xlab("Outcome measure")
@
\caption{Frequencies of some outcome measures for the effects in the dataset. 5593 measures with other outcome measures are excluded}
\label{study.outcomes}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
madata %>% group_by(file.nr) %>% distinct(study.name) %>% count() %>%
  filter(n < 50) %>%
  full_join( madata %>% group_by(file.nr) %>% distinct(study.name) %>% count %>%
               filter(n > 49) %>% mutate(n = 50)) %>%
  group_by(n) %>% count() %>%
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() + xlab("Study number") + ylab("Reviews Frequency")
@
\caption{Empirical distribution of number of studies per review}
\label{studies.per.review}
\end{figure}

\begin{figure}
<<echo=FALSE>>=
madata %>% group_by(file.nr) %>% count() %>%
  filter(n < 50) %>%
  full_join( madata %>% group_by(file.nr) %>% distinct(comparison.name) %>% count %>%
               filter(n > 49) %>% mutate(n = 50)) %>%
  group_by(n) %>% count() %>%
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() +  xlab("Comparisons") + ylab("Review Frequency")
# mean.diff.out <- madata %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
#   ungroup() %>% group_by(file.nr) %>% 
#   count() %>% ungroup %>% summarise(mean = mean(n))
# 
# median.diff.out <- madata %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
#   ungroup() %>% group_by(file.nr) %>% 
#   count() %>% ungroup %>% summarise(mean = median(n))
@
\caption{Empirical distribution of number of comparisons per review}
\label{comparisons.per.review}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
madata %>% group_by(file.nr) %>% distinct(comparison.name) %>% count() %>%
  filter(n < 15) %>%
  full_join( madata %>% group_by(file.nr) %>% distinct(comparison.name) %>% count %>%
               filter(n > 14) %>% mutate(n = 15)) %>%
  group_by(n) %>% count() %>%
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() +  xlab("Comparison type number") + ylab("Review frequency")

mean.diff.out <- madata %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
  ungroup() %>% group_by(file.nr) %>% 
  count() %>% ungroup %>% summarise(mean = mean(n))

median.diff.out <- madata %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
  ungroup() %>% group_by(file.nr) %>% 
  count() %>% ungroup %>% summarise(mean = median(n))
@
\caption{Empirical distribution of number of different research subjects per review}
\label{subjects.per.review}
\end{figure}

@@

<<echo=FALSE, results='asis'>>=
cum.repr.trials.nosub <- yodata %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
  filter(n < 15) %>%
  full_join( yodata %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
               filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>% 
  ungroup() %>% arrange(desc(n)) %>% mutate(csum  = cumsum(nn)) %>% arrange(n)

cum.repr.trials.subg <- yodata %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  filter(n < 15) %>%
  full_join( yodata %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
               filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>% 
  ungroup() %>% arrange(desc(n)) %>% mutate(csum  = cumsum(nn)) %>% arrange(n)

>>>>

<<echo=FALSE>>=
# PATH_HOME = path.expand("~") # user home
# PATH = file.path(PATH_HOME, 'thesis/code')
# source(file.path(PATH, 'prepare.R'))

PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'Data/PubBias')
FILE = 'cochrane_2018-06-09.csv'
PATH_DATA = file.path(PATH, 'data')
PATH_CODE = file.path(PATH, 'code')
PATH_RESULTS = file.path(PATH, 'results')
PATH_FIGURES = file.path(PATH_RESULTS, 'figures')

file_results = "pb.RData"

source(file.path(PATH_CODE, 'PubBias_functions.R'))

data = pb.readData(path = PATH_DATA, file = FILE)
tmp = pb.clean(data)
data = tmp[[1]]
aliases = tmp[[2]]

@@

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8, fig.height=3,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

@@

\subsection{Structure and Content}
The dataset consists of \Sexpr{length(unique(data$file.nr))} systematic reviews from the Cochrane Library with \Sexpr{length(unique(data$study.name))} studies.
Each study provides data of (multiple) comparisons of clinical interventions. 
% It can not be ruled out that some comparisons are retrospective, e.g. from observational studies. 
In Table \ref{barbiturate.row}, two comparisons from a systematic review about effects of barbiturates are shown as they are given in the dataset. As can be seen, the comparison is further specified by the variables in the columns. One row of the dataset is one comparison.

<<echo=FALSE, results = 'asis'>>=
tp1 <- arrange(filter(data, file.nr == 21) %>% 
          select(Study = study.name, Comparison_type = comparison.name, Outcome = outcome.name, 
                 Events = events1, Total = total1, Events_c = events2, Total_c = total2) %>% 
            slice(c(1,3)), Study)
print(xtable(tp1, label = "barbiturate.row", caption = "Example of two comparisons as given in the dataset. Events denotes the count of events in the treatment group while Events c the count of events in the group compared to. Further descriptive variables have been ommitted"), include.rownames = F, size = "scriptsize")
@

@@

<<echo=FALSE, results='asis'>>=
tp1 <- arrange(filter(data, file.nr == 21) %>% 
          select(Study = study.name, Comparison = comparison.name,Outcome = outcome.name), Study)
tp1[tp1$Study == "P\303\251rez-B\303\241rcena 2008", 1] = "Perez-Barcena 2008"
print(xtable(tp1, label = "barbiturates", caption = "Barbiturate and head injury review. In the columns, study names, comparison types and outcome measure of the comparisons are given"), include.rownames = F, size = "footnotesize")
@

@@

\vspace{0mm}
Having provided an overview over the dataset, now, some more specific information is provided. The dataset consists of \Sexpr{dim(data)[1]} comparisons and has \Sexpr{dim(data)[2]} variables that specify the comparisons. Information about missing values in the dataset is given in Table \ref{missing}. For variables as research subject, outcome and subgroup name and event counts there are no missing values. The relative amount of missing values is very low except for study years.

<<echo=FALSE>>=
missing.mean <- data %>% filter(!is.na(sd1) & sd1 > 0) %>% filter(!is.na(sd2) & sd2 > 0) %>% 
  summarise(total = length(sd1), na  = sum(is.na(mean1) | is.na(mean1))) %>% select(na)

missing.sd <- data %>% filter(!is.na(mean1) & !is.na(mean2)) %>% filter(mean1 != 0 & mean2 != 0) %>%
  summarise(total = length(mean1), na.sd  = sum(is.na(sd1) | is.na(sd1)), null.sd = length(sd1[sd1 <= 0 | sd2 <= 0])) %>% 
  mutate(miss = na.sd + null.sd) %>% select(miss)

missing.ssize <- data %>% summarize(total = length(total1), null.ss = length(total1[total1 <= 0 | total2 <= 0])) %>% 
  select(null.ss)

missing.effect <- sum(is.na(data$effect))

missing.year <- sum(is.na(data$study.year)) + length(data$study.year[data$study.year == 0])

# missing.outcome <- sum(is.na(data$effect))

# null.events <- data %>% filter(is.na(mean1) | mean1 == 0) %>% filter(is.na(mean2) | mean2 == 0) %>% 
#   filter(is.na(sd1) | sd1 <= 0) %>% filter(is.na(sd2) | sd2 <= 0) %>% 
#   summarise(null.ev = length(total1[events1 <= 0 & events2 <= 0]), total = length(events1)) %>% 
#   mutate(f = null.ev/total) %>% select(f)
# 
# null.events.single <- data %>% filter(is.na(mean1) | mean1 == 0) %>% filter(is.na(mean2) | mean2 == 0) %>% 
#   filter(is.na(sd1) | sd1 <= 0) %>% filter(is.na(sd2) | sd2 <= 0) %>% 
#   summarise(null.ev = length(total1[events1 <= 0 | events2 <= 0]), total = length(events1)) %>% 
#   mutate(f = null.ev/total) %>% select(f)
# 
# missing.events <- data %>% filter(is.na(mean1) | mean1 == 0) %>% filter(is.na(mean2) | mean2 == 0) %>% 
#   filter(is.na(sd1) | sd1 <= 0) %>% filter(is.na(sd2) | sd2 <= 0) %>% 
#   summarise(na.ev = sum(is.na(events1) | is.na(events2)), total = length(events1)) %>% 
#   mutate(f = na.ev/total) %>% select(f)
@

@@

More properties of the reviews, the studies and the comparisons in the dataset will be provided on the following pages. The publication dates of the studies included in the dataset are shown in Figure \ref{study.years}. Most studies were published after 2000.

\begin{figure}
<<echo=FALSE>>=
data %>% filter(study.year < 2019 & study.year > 1920) %>% ggplot(aes(x = study.year)) + geom_histogram(col = "gray15", fill = "dodgerblue") +
  theme_bw() + xlab("Year") + ylab("Study frequency")
@
\caption{Frequencies of study publication years in the dataset. 44655 were excluded due to likely wrong indications}
\label{study.years}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
data %>% group_by(outcome.measure.new) %>% count() %>% arrange(desc(n)) %>% ungroup() %>% filter(row_number() < 9) %>% 
  ggplot(aes(x = outcome.measure.new, y = n)) +
  geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() + ylab("Comparison frequency") + xlab("Outcome measure")
@
\caption{Frequencies of some outcome measures for the effects in the dataset. 5593 measures with other outcome measures are excluded}
\label{study.outcomes}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
data %>% group_by(file.nr) %>% distinct(study.name) %>% count() %>%
  filter(n < 50) %>%
  full_join( data %>% group_by(file.nr) %>% distinct(study.name) %>% count %>%
               filter(n > 49) %>% mutate(n = 50)) %>%
  group_by(n) %>% count() %>%
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() + xlab("Study number") + ylab("Reviews Frequency")
@
\caption{Empirical distribution of number of studies per review}
\label{studies.per.review}
\end{figure}

\begin{figure}
<<echo=FALSE>>=
data %>% group_by(file.nr) %>% count() %>%
  filter(n < 50) %>%
  full_join( data %>% group_by(file.nr) %>% distinct(comparison.name) %>% count %>%
               filter(n > 49) %>% mutate(n = 50)) %>%
  group_by(n) %>% count() %>%
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() +  xlab("Comparisons") + ylab("Review Frequency")
# mean.diff.out <- data %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
#   ungroup() %>% group_by(file.nr) %>% 
#   count() %>% ungroup %>% summarise(mean = mean(n))
# 
# median.diff.out <- data %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
#   ungroup() %>% group_by(file.nr) %>% 
#   count() %>% ungroup %>% summarise(mean = median(n))
@
\caption{Empirical distribution of number of comparisons per review}
\label{comparisons.per.review}
\end{figure}

@@

\begin{figure}
<<echo=FALSE>>=
data %>% group_by(file.nr) %>% distinct(comparison.name) %>% count() %>%
  filter(n < 15) %>%
  full_join( data %>% group_by(file.nr) %>% distinct(comparison.name) %>% count %>%
               filter(n > 14) %>% mutate(n = 15)) %>%
  group_by(n) %>% count() %>%
  ggplot(aes(x = n, y = nn)) + geom_col(col = "gray15", fill = "dodgerblue") +
  theme_bw() +  xlab("Comparison type number") + ylab("Review frequency")

mean.diff.out <- data %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
  ungroup() %>% group_by(file.nr) %>% 
  count() %>% ungroup %>% summarise(mean = mean(n))

median.diff.out <- data %>% group_by(file.nr, comparison.nr) %>% distinct(outcome.name) %>% 
  ungroup() %>% group_by(file.nr) %>% 
  count() %>% ungroup %>% summarise(mean = median(n))
@
\caption{Empirical distribution of number of different comparison types per review}
\label{subjects.per.review}
\end{figure}

@@

<<echo=FALSE, results='asis'>>=
cum.repr.trials.nosub <- data %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
  filter(n < 15) %>%
  full_join( data %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
               filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>% 
  ungroup() %>% arrange(desc(n)) %>% mutate(csum  = cumsum(nn)) %>% arrange(n)

cum.repr.trials.subg <- data %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  filter(n < 15) %>%
  full_join( data %>% group_by(file.nr, comparison.nr, outcome.nr) %>% count %>% group_by(n) %>% count %>%
               filter(n > 14) %>% ungroup %>% summarise(n = 15, nn = sum(nn))) %>% 
  ungroup() %>% arrange(desc(n)) %>% mutate(csum  = cumsum(nn)) %>% arrange(n)
